Work Fragmentation in Developer Interaction Data

### 2016-03-18 paper_2.tex
text on RQ2

In this section we analyze whether or not the duration of the interruptions is a factor in the relationship between interruptions and developer productivity. Under the hypothesis that prolonged interruptions require more time to recover, we built two groups of sessions based on the proportion of long interruptions. As we described previously we consider that a prolonged interruption has a duration of $\geq$ 12 minutes, according to previous work \cite{GM04, KaptelininN07}. Prolonged interruptions could be due to an extended interruption, such as a coffee break or a meeting.

@@

Following the conclusions on RQ1, the number of interruptions has an impact on the metrics and is a variable that should not be ignored. So, we split every group in accordance to the number of interruptions: this time, instead of taking the quantiles of the distribution to create groups we created a group for every number of interruptions up to seven. Since the results in RQ1 show that as the number of interruptions increases the difference between the groups minimizes, we believe that having more distinctive groups instead of groups created by ranges is also valid. Thus, for the three metrics, we created the groups of sessions according to the number on interruptions, between one and seven, and then we split every group in low and high, according to the proportion of prolonged interruptions.

The Figure \ref{fig:dur_mylyn} shows the median of the metrics when grouping the sessions by the number of interruptions and the proportion of prolonged interruptions with Mylyn, and the Figure \ref{fig:dur_udc} shows the same with UDC. According to our theories, it is better to have a low number of interruptions and a low proportion of prolonged interruptions to have higher rates of productivity, and the results agree with this statement. First, we observe that (as we conclude in RQ1) when the number of interruption increases the median of the metric decreases until reaching a point (approximately between three and four interruptions) where the change is minimal and continues that way for the rest of the groups. Second, almost in every group the median value of groups with low proportion of prolonged interruptions is greater than in groups with a high proportion. The latter observation is more clear in the edits per minute metric in both datasets, for we can see that at some points the medians intertwine in the selections per minutes and edit ratio, specially with Mylyn.

When having a small number of interruptions (between 1 and 3) the difference between sessions with a low or high proportion of prolonged interruptions is more noticeable, and the group of sessions with a low proportion has higher median values. As the number of interruptions increases the differences get smaller, but we observe that the group of low proportion of prolonged interruptions always shows bigger rates. This comments apply on the edits and selections per minute, and to a lesser extent to the edit ratio, where the changes between groups (both by number of interruptions and proportion) are small.

@@

When comparing the results between the datasets, we can see the same observations described before. Therefore, we can conclude that the duration of the interruption has an impact when having a low number of interruptions, but when the number is bigger the impact is very subtle. We tested the statistical significance of this results and show the results in the Table \ref{tbl:effect_size_dur_mylyn} for Mylyn, and Table \ref{tbl:effect_size_dur_udc} for UDC. For every group according to the number of interruptions, we measured the effect between the low and high groups of sessions, with the Cohen's $d$ test and U-test. The groups with one to three interruptions have moderate size effects, in comparison with the other groups with more interruptions that have lower size effects; this matches our observations via visual inspection on a more defined difference between the low and high groups when having a small number of interruptions. Starting from (approximately) three interruptions, the distributions of the metrics are very alike, except in some cases like the edit ratio with seven interruptions in Mylyn, where the median is almost the same but the distribution of the values are different enough to have a medium effect size. The bold numbers are those with a significant effect size.

>>>>

In this section we analyze whether or not the duration of interruptions is a factor in the relationship between interruptions and developer productivity. Under the hypothesis that prolonged interruptions require more time to recover from, we built two groups of sessions based on the proportion of long interruptions. As we described previously we consider that a prolonged interruption has a duration of $\geq$ 12 minutes---a decision based on the literature\cite{GM04, KaptelininN07}. %Prolonged interruptions could be due to an extended interruption, such as a coffee break or a meeting.

@@

Following the conclusions on RQ1, the number of interruptions has an impact on the metrics and should not be ignored. Further, as the number of interruptions increases the difference between the groups minimizes.  %we believe that having   more distinctive groups instead of groups created by ranges is also valid.
 Thus we want to compare the proportions of prolonged interruptions among sessions that have the \emph{same} number of interruptions. We hence split the \emph{low} and \emph{high} groups in subgroups of sessions which have the same number of interruptions. We consider groups of up to seven sessions, as larger groups are small. We finally compare pairs of \emph{low} and \emph{high} groups that have the same number of interruptions, expecting higher productivity indicators in the \emph{low} groups.

%We split every group in accordance to the number of interruptions: this time, instead of taking the quantiles of the distribution to create groups we created a group for every number of interruptions up to seven. Since the results in RQ1 show that as the number of interruptions increases the difference between the groups minimizes, we believe that having more distinctive groups instead of groups created by ranges is also valid. Thus, for the three metrics, we created the groups of sessions according to the number on interruptions, between one and seven, and then we split every group in low and high, according to the proportion of prolonged interruptions.

The Figure \ref{fig:dur_mylyn} shows the median of the metrics when grouping the sessions by the number of interruptions and the proportion of prolonged interruptions with Mylyn, and the Figure \ref{fig:dur_udc} shows the same with UDC. According to our hypothesis, we expect sessions that have a low number of interruptions and a low proportion of prolonged interruptions to have higher rates of productivity, and the results mostly agree with this statement. First, we observe that (as we conclude in RQ1) when the number of interruption increases the median of the metric decreases until reaching a point (approximately between three and four interruptions) where the change is minimal and continues that way for the rest of the groups. Second, in the vast majority of groups the median value of the \emph{low} group is greater than the \emph{high} group---this happens 35 times, while the opposite is true twice, and the median are equals five times. This is the clearest in the edits per minute metric in both datasets, as we can see that at some points the medians intertwine in the selections per minutes and edit ratio, especially in Mylyn.

When having a small number of interruptions (between 1 and 3) the difference between sessions with a low or high proportion of prolonged interruptions is very noticeable, and the group of sessions with a low proportion has higher median values. As the number of interruptions increases the differences get smaller, but we observe that the group of low proportion of prolonged interruptions usually shows higger values. This applies on the edits and selections per minute, and to a lesser extent to the edit ratio, where the changes between groups (both by number of interruptions and proportion) are comparatively smaller.

Since these results apply to both datasets, we can conclude that we observe a relationship between the duration of the interruptions and the productivity, with a caveat: the relation is much more pronounced when the amount of interruptions is lower; as the number of interruptions grows it becomes very subtle. We speculate, but can not confirm, that this is due to the noise in the dataset, as sessions with more interruptions are usually longer and thus may exhibit more variation due to other factors.

%When comparing the results between the datasets, we can see the same observations described before. Therefore, we can conclude that the duration of the interruption has an impact when having a low number of interruptions, but when the number is bigger the impact is very subtle.

@@

We tested the statistical significance of this results and show the results in the Table \ref{tbl:effect_size_dur_mylyn} for Mylyn, and Table \ref{tbl:effect_size_dur_udc} for UDC. For every group according to the number of interruptions, we measured the effect between the low and high groups of sessions, with the Cohen's $d$ test and U-test. The groups with one to three interruptions have moderate size effects, in comparison with the other groups with more interruptions that have lower size effects; this matches our observations via visual inspection on a more defined difference between the low and high groups when having a small number of interruptions. Starting from (approximately) three interruptions, the distributions of the metrics are very alike, except in some cases like the edit ratio with seven interruptions in Mylyn, where the median is almost the same but the distribution of the values are different enough to have a medium effect size. The bold numbers are those with a significant effect size.
